---
title: "Mapping SLCO Preschools"
author: "Gwendolyn Reynolds "
date: "7/6/2017"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}

library(tigris)
library(acs)
library(stringr)
library(dplyr)
library(leaflet)
library(ggplot2)
library(tidyr)
library(ggmap)
library(spatialEco)


knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(dpi = 150)
knitr::opts_chunk$set(message = T) #Use this to suppress all messages from chunks for finalized report
knitr::opts_chunk$set(warning = T) #Use this to suppress all warnings from chunks for finalized report

# Sorenson Impact colors we use in ggplot
SI_design <- list(color_1 = "#005060", color_medium = "#1A818F", color_light = "#25AFC1", color_grey = "#58585A", color_light_grey = "E6E7E8")

# This will fill bar colors with Sorenson Impact colors and default to theme_minimal
update_geom_defaults("bar", list(fill = SI_design$color_1))
theme_set(theme_minimal())

setwd("~/Google Drive/SI/DataScience/")
knitr::opts_knit$set(root.dir = '~/Google Drive/SI/DataScience/SLCPreschoolStudy2017/')

```

```{r importing the addresses, echo=FALSE}

 origAddress <- read.csv("./raw_data/ProviderAddress.csv", stringsAsFactors = FALSE)
# 
# ## format the address so the streets are with the city and state
# 
# origAddress <- origAddress %>%
#   unite(addresses, Address, City, State, sep = ", ")
# 
# # Initialize the data frame
# geocoded <- data.frame(stringAsFactors = FALSE)
# 
# ##have to trim the white space because there is something funky going on when i read in the csv file. FUCK!
# origAddress <- origAddress %>%
#  mutate(addresses = trimws(addresses)) %>% 
#  mutate_geocode(addresses) 
# 
# origAddress <- origAddress %>%
#   select(Business, addresses, Zip.Code, Care.Type, lon, lat)
# 
# origAddress <- origAddress %>%
#   mutate(Business = ifelse(Business=="", Care.Type, Business))
# 
# origAddress <- origAddress %>%
#   rename(long = lon)
# 
# write.csv(origAddress, "./raw_data/latlongaddress.csv")

origAddress <- read.csv("./latlongaddress.csv", stringsAsFactors = FALSE)

```

```{r, echo=FALSE}

# qualaddress <- read.csv("./raw_data/NAEYCPreschools.csv", stringsAsFactors = FALSE)
# 
# qualaddress <- qualaddress %>% 
#   unite(addresses, Address, City, State, sep = ", ")
# 
# geocoded <- data.frame(stringsAsFactors = FALSE)
# 
# qualaddress <- qualaddress %>%
#   filter(Program.Name!="Adele and Dale Young Child Development Laboratory")
# 
# qualaddress <- qualaddress %>%
#   mutate(addresses = ifelse(addresses=="Bldg. 5124 Kister Avenue, Dugway, Utah", "5124 Kister Avenue, Dugway, Utah", addresses))
# 
# read.csv("./raw_data/quallatlongaddress.csv", stringsAsFactors = FALSE)
# 
# # Loop through the addresses to get the latitude and longitude of each address and add it to the
# # # origAddress data frame in new columns lat and lon
#  for(i in 1:nrow(qualaddress))
#  {
#    #print("working..")
#    result <- geocode(qualaddress$addresses[i], output = "latlona", source = "google")
#   qualaddress$lon[i] <- as.numeric(result[1])
#   qualaddress$lat[i] <- as.numeric(result[2])
#   qualaddress$geoAddress[i] <- as.character(result[3])
# }
# # # 
# origAddress %>%
#   filter(is.na(lat)) %>%
#   tally()
# # # 
# qualaddress <- qualaddress %>%
#    select(Program.Name, addresses, lon, lat, geoAddress)
# 
# # qualaddress <- qualaddress %>%
# #   mutate(Program.Name = ifelse(Business=="", Care.Type, Business))
# 
#  qualaddress <- qualaddress %>%
#    rename(long = lon)

# write.csv(qualaddress, "./raw_data/quallatlongaddress.csv")

 qualaddress <- read.csv("./raw_data/quallatlongaddress.csv", stringsAsFactors = FALSE)



```

## Head Start Locations in Utah
The map below shows all Head Start locations in the state of Utah. 

```{r map head start, echo=FALSE}

headstart <- read.csv("./raw_data/HeadStartUtah.csv", stringsAsFactors = FALSE)

headstart <- headstart %>% 
  rename(long = longitude) %>% 
  rename(lat = latitude)


map20<-leaflet(headstart) %>%
  addMarkers(~long, ~lat, popup = ~as.character(name), label = ~as.character(name)) %>% 
  addProviderTiles("CartoDB.Positron") 

map20

```

## NAEYC Certified Preschools
The map below shows all NAEYC Certified preschools in the state of Utah. They are predominantly located in the Salt Lake valley. 

```{r}

map10<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") 

map10

```

## All Childcare Providers From Website
The following map shows all of the childcare providers listed at the website: 

```{r all childcare, echo=FALSE, warning=FALSE, message=FALSE}

map100<-leaflet(origAddress) %>%
  addProviderTiles("CartoDB.Positron") %>% 
  addMarkers(clusterOptions = markerClusterOptions()) 

map100

```



```{r, echo=FALSE, message=FALSE, warning=FALSE}

##must find the fips number for the counties i am including. i only want to include sl county and i want to map by school district 

counties <- c(035)
schooldistrictunified <- school_districts(state = 'UT', type = "unified")
```

## Total Population in SLCO School Districts
We are able to use the American Community Survey, which is run by the United States Census Bureau, to map the total population in Salt Lake County School Districts. The ACS was last updated in 2015, so population counts are as of 2015. We then overlay the NAEYC certified preschools on top of total population. 

```{r get tabular data, echo=FALSE, message=FALSE, warning=FALSE}

# install api key using key you get from census website
api.key.install(key="0cdac4dbbd32b4e4874b79ce6e8fee07d12a7b3a")

#create a geographic set to grab tabular data 
geo <- geo.make(state=c("UT"), 
                school.district.unified ="*") 

#update endyear to 2015

population <- acs.fetch(endyear = 2015, span = 5, geography = geo, 
                     table.number = "B01003", col.names = "pretty")

#names(attributes(population))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"  
# [8] "estimate"       "standard.error" "class"    

#attr(population, "acs.colnames")

#convert to data.frame for merging
pop_df <- data.frame(paste0(str_pad(population@geography$state, 2, "left", pad="0"), 
                                #str_pad(population@geography$county, 3, "left", pad = "0"), 
                                str_pad(population@geography$schooldistrictunified, 5, "left", pad = "0")), 
                         population@estimate[,c("Total Population: Total")], 
                         stringsAsFactors = FALSE)

pop_df <- select(pop_df, 1:2)
rownames(pop_df)<-1:nrow(pop_df)
names(pop_df)<-c("GEOID", "totalpop")
#pop_df$percent <- 100*(pop_df$wkids/(poverty_df$didnottotal+poverty_df$didtotal))

##time to merge

pop_merged<- geo_join(schooldistrictunified, pop_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
pop_merged <- pop_merged[pop_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", pop_merged$NAME, "</b><br>", "Total Population ", prettyNum(round(pop_merged$totalpop,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = pop_merged$totalpop
)

map3<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = pop_merged, 
              fillColor = ~pal(totalpop), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = pop_merged$totalpop, 
            position = "bottomright", 
            title = "Total Population"#,
            #labFormat = labelFormat(prefix = " ")
           ) 
map3

```

## Total Population Under 5 Years Old in SLCO School Districts
We use the same ACS data (from 2015) to map the total population of children under 5 years old in the Salt Lake County School Districts. 

```{r population under 3, echo=FALSE, message=FALSE, warning=FALSE}


under3 <- acs.fetch(endyear = 2015, span = 5, geography = geo, 
                     table.number = "B09001", col.names = "pretty")

#names(attributes(under3))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"  
# [8] "estimate"       "standard.error" "class"    

#attr(under3, "acs.colnames")

#convert to data.frame for merging
popunder5_df <- data.frame(paste0(str_pad(under3@geography$state, 2, "left", pad="0"), 
                                #str_pad(population@geography$county, 3, "left", pad = "0"), 
                                str_pad(under3@geography$schooldistrictunified, 5, "left", pad = "0")), 
                         under3@estimate[,c("Population Under 18 Years by Age: Total:", "Population Under 18 Years by Age: In households:", "Population Under 18 Years by Age: In households: Under 3 years", "Population Under 18 Years by Age: In households: 3 and 4 years", "Population Under 18 Years by Age: In households: 5 years", "Population Under 18 Years by Age: In households: 6 to 8 years", "Population Under 18 Years by Age: In households: 9 to 11 years", "Population Under 18 Years by Age: In households: 12 to 14 years", "Population Under 18 Years by Age: In households: 15 to 17 years", "Population Under 18 Years by Age: In group quarters")], 
                         stringsAsFactors = FALSE)

popunder5_df <- select(popunder5_df, 1:6)
rownames(popunder5_df)<-1:nrow(popunder5_df)
names(popunder5_df)<-c("GEOID", "totalpopunder", "totpopunderhouse", "under3", "threeand4", "fiveyears")
popunder5_df$totalunder5 <- (popunder5_df$under3 + popunder5_df$threeand4 + popunder5_df$fiveyears)

##time to merge

popunder5_merged<- geo_join(schooldistrictunified, popunder5_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
popunder5_merged <- popunder5_merged[popunder5_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", popunder5_merged$NAME, "</b><br>", "Total Population Under 5: ", prettyNum(round(popunder5_merged$totalunder5,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = popunder5_merged$totalunder5
)

map2<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>%   
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = popunder5_merged, 
              fillColor = ~pal(totalunder5), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = popunder5_merged$totalunder5, 
            position = "bottomright", 
            title = "Total Population Under 5"#,
            #labFormat = labelFormat(prefix = " ")
           ) 
map2

```

## Percent of Population Under 5 Years Old in SLCO School Districts
We use the same ACS data (from 2015) to map the percent of the total population that are children under 5 years old in the Salt Lake County School Districts. 

```{r percent of population under 3, echo=FALSE, message=FALSE, warning=FALSE}


under3 <- acs.fetch(endyear = 2015, span = 5, geography = geo, 
                     table.number = "B09001", col.names = "pretty")

#names(attributes(under3))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"  
# [8] "estimate"       "standard.error" "class"    

#attr(under3, "acs.colnames")

#convert to data.frame for merging
popunder5_df <- data.frame(paste0(str_pad(under3@geography$state, 2, "left", pad="0"), 
                                #str_pad(population@geography$county, 3, "left", pad = "0"), 
                                str_pad(under3@geography$schooldistrictunified, 5, "left", pad = "0")), 
                         under3@estimate[,c("Population Under 18 Years by Age: Total:", "Population Under 18 Years by Age: In households:", "Population Under 18 Years by Age: In households: Under 3 years", "Population Under 18 Years by Age: In households: 3 and 4 years", "Population Under 18 Years by Age: In households: 5 years", "Population Under 18 Years by Age: In households: 6 to 8 years", "Population Under 18 Years by Age: In households: 9 to 11 years", "Population Under 18 Years by Age: In households: 12 to 14 years", "Population Under 18 Years by Age: In households: 15 to 17 years", "Population Under 18 Years by Age: In group quarters")], 
                         stringsAsFactors = FALSE)

popunder5_df <- select(popunder5_df, 1:6)
rownames(popunder5_df)<-1:nrow(popunder5_df)
names(popunder5_df)<-c("GEOID", "totalpopunder", "totpopunderhouse", "under3", "threeand4", "fiveyears")
popunder5_df$totalunder5 <- (100* (popunder5_df$under3 + popunder5_df$threeand4 + popunder5_df$fiveyears)/(pop_df$total))

##time to merge

popunder5_merged<- geo_join(schooldistrictunified, popunder5_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
popunder5_merged <- popunder5_merged[popunder5_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", popunder5_merged$NAME, "<br></b>", "Percent of Population Under 5: ", round(popunder5_merged$totalunder5,2))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = popunder5_merged$totalunder5
)

map1<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = popunder5_merged, 
              fillColor = ~pal(totalunder5), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = popunder5_merged$totalunder5, 
            position = "bottomright", 
            title = "Percent of Population Under 5",
            labFormat = labelFormat(suffix = "%")
           ) 
map1

```

## Total Population of 3 and 4 Year Olds in SLCO School Districts
We then zero in specifically on 3 and 4 year olds in Salt Lake County School Districts.

```{r population 3&4, echo=FALSE, message=FALSE, warning=FALSE}


under3 <- acs.fetch(endyear = 2015, span = 5, geography = geo, 
                     table.number = "B09001", col.names = "pretty")

#names(attributes(under3))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"  
# [8] "estimate"       "standard.error" "class"    

#attr(under3, "acs.colnames")

#convert to data.frame for merging
popunder5_df <- data.frame(paste0(str_pad(under3@geography$state, 2, "left", pad="0"), 
                                #str_pad(population@geography$county, 3, "left", pad = "0"), 
                                str_pad(under3@geography$schooldistrictunified, 5, "left", pad = "0")), 
                         under3@estimate[,c("Population Under 18 Years by Age: Total:", "Population Under 18 Years by Age: In households:", "Population Under 18 Years by Age: In households: Under 3 years", "Population Under 18 Years by Age: In households: 3 and 4 years", "Population Under 18 Years by Age: In households: 5 years", "Population Under 18 Years by Age: In households: 6 to 8 years", "Population Under 18 Years by Age: In households: 9 to 11 years", "Population Under 18 Years by Age: In households: 12 to 14 years", "Population Under 18 Years by Age: In households: 15 to 17 years", "Population Under 18 Years by Age: In group quarters")], 
                         stringsAsFactors = FALSE)

popunder5_df <- select(popunder5_df, 1:6)
rownames(popunder5_df)<-1:nrow(popunder5_df)
names(popunder5_df)<-c("GEOID", "totalpopunder", "totpopunderhouse", "under3", "threeand4", "fiveyears")

##time to merge

popunder5_merged<- geo_join(schooldistrictunified, popunder5_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
popunder5_merged <- popunder5_merged[popunder5_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", popunder5_merged$NAME, "</b><br>", "Total Population of 3 and 4 Year Olds: ", prettyNum(round(popunder5_merged$threeand4,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = popunder5_merged$threeand4
)

map4<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = popunder5_merged, 
              fillColor = ~pal(threeand4), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = popunder5_merged$threeand4, 
            position = "bottomright", 
            title = "Total Population 3 & 4 Year Olds"#,
            #labFormat = labelFormat(prefix = " ")
           ) 
map4

```

## Percent of Population that's 3 and 4 Years Old in SLCO School Districts
We then map the percent of the total population that are 3 and 4 years old in Salt Lake County School Districts.

```{r percent population 3&4, echo=FALSE, message=FALSE, warning=FALSE}


under3 <- acs.fetch(endyear = 2015, span = 5, geography = geo, 
                     table.number = "B09001", col.names = "pretty")

#names(attributes(under3))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"  
# [8] "estimate"       "standard.error" "class"    

#attr(under3, "acs.colnames")

#convert to data.frame for merging
popunder5_df <- data.frame(paste0(str_pad(under3@geography$state, 2, "left", pad="0"), 
                                #str_pad(population@geography$county, 3, "left", pad = "0"), 
                                str_pad(under3@geography$schooldistrictunified, 5, "left", pad = "0")), 
                         under3@estimate[,c("Population Under 18 Years by Age: Total:", "Population Under 18 Years by Age: In households:", "Population Under 18 Years by Age: In households: Under 3 years", "Population Under 18 Years by Age: In households: 3 and 4 years", "Population Under 18 Years by Age: In households: 5 years", "Population Under 18 Years by Age: In households: 6 to 8 years", "Population Under 18 Years by Age: In households: 9 to 11 years", "Population Under 18 Years by Age: In households: 12 to 14 years", "Population Under 18 Years by Age: In households: 15 to 17 years", "Population Under 18 Years by Age: In group quarters")], 
                         stringsAsFactors = FALSE)

popunder5_df <- select(popunder5_df, 1:6)
rownames(popunder5_df)<-1:nrow(popunder5_df)
names(popunder5_df)<-c("GEOID", "totalpopunder", "totpopunderhouse", "under3", "threeand4", "fiveyears")
popunder5_df$percent <- (100* (popunder5_df$threeand4/pop_df$totalpop))

##time to merge

popunder5_merged<- geo_join(schooldistrictunified, popunder5_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
popunder5_merged <- popunder5_merged[popunder5_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", popunder5_merged$NAME, "</b><br>", "Percent of Population 3 and 4 Years Old: ", prettyNum(round(popunder5_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = popunder5_merged$percent
)

map5<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = popunder5_merged, 
              fillColor = ~pal(percent), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = popunder5_merged$percent, 
            position = "bottomright", 
            title = "Percent of Population 3 & 4 Years Old",
            labFormat = labelFormat(suffix = "%")
           ) 
map5

```

## Childcare Subsidies in Salt Lake County School Districts
We used the data we received on childcare to map the total amount of childcare subsidies given by Salt Lake County School Districts.  

```{r subsidies by provider and child, echo=FALSE}

# subsidy<- read.csv("./latlongaddress.csv", stringsAsFactors = FALSE)
# 
# subsidy <- subsidy %>% 
#   na.exclude(Subsidized.Amt)
#  
#  subsidypoints <- subsidy %>% 
#    select(lon, lat) 
# 
# xy <- subsidypoints[,c(1, 2)]
# 
# spdf <- SpatialPointsDataFrame(coords = xy, data = subsidypoints,
#                                proj4string = CRS("+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0 "))
# 
#  new_df <- sp::over(spdf, schooldistrictunified) 
#  
#  subsidy$district <- new_df$NAME
#  subsidy$GEOID <- new_df$GEOID
# 
# districtsubsidy <-  subsidy %>% 
#   group_by(GEOID, district) %>% 
#   summarize(totalsubsidy = sum(Subsidized.Amt, na.rm=TRUE))
# 
# write.csv(districtsubsidy, "./districtsubsidy.csv")

districtsubsidy <- read.csv("~/Github/ACSMapping/districtsubsidy.csv", stringsAsFactors = FALSE)
   
subsidy_merged<- geo_join(schooldistrictunified, districtsubsidy, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
subsidy_merged <- subsidy_merged[subsidy_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", subsidy_merged$NAME, "</b><br>", "Total Childcare Subsidy by District: ", prettyNum(round(subsidy_merged$totalsubsidy,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = subsidy_merged$totalsubsidy
)

map33<-leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = subsidy_merged, 
              fillColor = ~pal(totalsubsidy), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = subsidy_merged$totalsubsidy, 
            position = "bottomright", 
            title = "Total Childcare Subsidy by District"#,
            #labFormat = labelFormat(prefix = " ")
           ) 
map33

```

## Childcare Subsidies in Salt Lake County School Districts Per Child Under 5
We used the data we received on childcare to map the amount of childcare subsidies per child under 5 in Salt Lake County School Districts.  

```{r subsidies by child, echo=FALSE}


subsidy_merged$percent <- (subsidy_merged$totalsubsidy/(popunder5_merged$under3 + popunder5_merged$threeand4 + popunder5_merged$fiveyears))

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", subsidy_merged$NAME, "</b><br>", "Childcare Subsidy Per Population Under 5: ", prettyNum(round(subsidy_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = subsidy_merged$percent
)

map37<-leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = subsidy_merged, 
              fillColor = ~pal(percent), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = subsidy_merged$percent, 
            position = "bottomright", 
            title = "Childcare Subsidy Per Population Under 5"#,
            #labFormat = labelFormat(prefix = "%")
           ) 
map37

```

## Median Household Income in Salt Lake County School Districts 
We used ACS data (from 2015) to map the median household income for each Salt Lake County school district.  


```{r median income, echo=FALSE, message=FALSE, warning=FALSE}


income <- acs.fetch(endyear = 2015, span = 5, geography = geo, 
                     table.number = "B19013", col.names = "pretty")

#names(attributes(income))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"  
# [8] "estimate"       "standard.error" "class"    

#attr(income, "acs.colnames")

#convert to data.frame for merging
income_df <- data.frame(paste0(str_pad(income@geography$state, 2, "left", pad="0"), 
                                #str_pad(population@geography$county, 3, "left", pad = "0"), 
                                str_pad(income@geography$schooldistrictunified, 5, "left", pad = "0")), 
                         income@estimate[,c("B19013. Median Household Income in the Past 12 Months (in 2015 Inflation-Adjusted Dollars): Median household income in the past 12 months (in 2015 Inflation-adjusted dollars)")],
                         stringsAsFactors = FALSE)

income_df <- select(income_df, 1:2)
rownames(income_df)<-1:nrow(income_df)
names(income_df)<-c("GEOID", "medianincome")
#income_df$percent <- (100* (popunder5_df$threeand4/pop_df$totalpop))

##time to merge

income_merged<- geo_join(schooldistrictunified, income_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
income_merged <- income_merged[income_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", income_merged$NAME, "</b><br>", "Median Household Income: ", prettyNum(round(income_merged$medianincome,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = income_merged$medianincome
)

map6 <- leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = income_merged, 
              fillColor = ~pal(medianincome), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = income_merged$medianincome, 
            position = "bottomright", 
            title = "Median Household Income"#,
            #labFormat = labelFormat(suffix = "%")
           ) 
map6

```

## Race: White Alone in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of people who marked their race/ethnicty as White: Alone for each Salt Lake County school district. 

```{r white people, echo=FALSE, message=FALSE, warning=FALSE}


race <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "C03002", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(race, "acs.colnames")

#convert to data.frame for merging
race_df <- data.frame(paste0(str_pad(race@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(race@geography$schooldistrictunified, 5, "left", pad = "0")),
                         race@estimate[,c("Hispanic or Latino by Race: Total:", "Hispanic or Latino by Race: Not Hispanic or Latino: White alone", "Hispanic or Latino by Race: Hispanic or Latino")],
                         stringsAsFactors = FALSE)

race_df <- select(race_df, 1:4)
rownames(race_df)<-1:nrow(race_df)
names(race_df)<-c("GEOID", "totalpop", "whitealone", "latino")
race_df$percent <- (100* (race_df$whitealone/race_df$totalpop))

##time to merge

race_merged<- geo_join(schooldistrictunified, race_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
race_merged <- race_merged[race_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", race_merged$NAME, "</b><br>", "Percent White (Alone): ", prettyNum(round(race_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = race_merged$percent
)

map7<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = race_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = race_merged$percent,
            position = "bottomright",
            title = "Percent White Alone",
            labFormat = labelFormat(suffix = "%")
           )
map7

```

## Race: Hispanic/Latino in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of people who marked their race/ethnicty as Hispanci/Latino for each Salt Lake County school district. 


```{r hispanic/latino people, echo=FALSE, message=FALSE, warning=FALSE}


race <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "C03002", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(race, "acs.colnames")

#convert to data.frame for merging
race_df <- data.frame(paste0(str_pad(race@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(race@geography$schooldistrictunified, 5, "left", pad = "0")),
                         race@estimate[,c("Hispanic or Latino by Race: Total:", "Hispanic or Latino by Race: Not Hispanic or Latino: White alone", "Hispanic or Latino by Race: Hispanic or Latino")],
                         stringsAsFactors = FALSE)

race_df <- select(race_df, 1:4)
rownames(race_df)<-1:nrow(race_df)
names(race_df)<-c("GEOID", "totalpop", "whitealone", "latino")
race_df$percent <- (100* (race_df$latino/race_df$totalpop))

##time to merge

race_merged<- geo_join(schooldistrictunified, race_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
race_merged <- race_merged[race_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", race_merged$NAME, "</b><br>", "Percent Hispanic/Latino: ", prettyNum(round(race_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = race_merged$percent
)

map8<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = race_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = race_merged$percent,
            position = "bottomright",
            title = "Percent Hispanic/Latino",
            labFormat = labelFormat(suffix = "%")
           )
map8

```


## Foreign Born in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of people who are foreign born (as opposed to native born) for each Salt Lake County school district. 

```{r foreign born, echo=FALSE}

foreign <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "B05012", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(foreign, "acs.colnames")

#convert to data.frame for merging
foreign_df <- data.frame(paste0(str_pad(foreign@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(foreign@geography$schooldistrictunified, 5, "left", pad = "0")),
                         foreign@estimate[,c("Nativity in the United States: Total:", "Nativity in the United States: Native", "Nativity in the United States: Foreign-Born")],
                         stringsAsFactors = FALSE)

foreign_df <- select(foreign_df, 1:4)
rownames(foreign_df)<-1:nrow(foreign_df)
names(foreign_df)<-c("GEOID", "totalpop", "native", "foreign")
foreign_df$foreignpercent <- (100* (foreign_df$foreign/foreign_df$totalpop))

##time to merge

foreign_merged<- geo_join(schooldistrictunified, foreign_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
foreign_merged <- foreign_merged[foreign_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", foreign_merged$NAME, "</b><br>", "Percent Foreign Born: ", prettyNum(round(foreign_merged$foreignpercent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = foreign_merged$foreignpercent
)

map11<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = foreign_merged,
              fillColor = ~pal(foreignpercent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = foreign_merged$foreignpercent,
            position = "bottomright",
            title = "Percent Foreign Born",
            labFormat = labelFormat(suffix = "%")
           )
map11

```

## Native Born in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of people who are native born (as opposed to foreign born) for each Salt Lake County school district. 

```{r percent not foreign born, echo=FALSE}

foreign_df$nativepercent <- (100* (foreign_df$native/foreign_df$totalpop))

##time to merge

foreign_merged<- geo_join(schooldistrictunified, foreign_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
foreign_merged <- foreign_merged[foreign_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", foreign_merged$NAME, "</b><br>", "Percent Native Born: ", prettyNum(round(foreign_merged$nativepercent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = foreign_merged$nativepercent
)

map12<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = foreign_merged,
              fillColor = ~pal(nativepercent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = foreign_merged$nativepercent,
            position = "bottomright",
            title = "Percent Native Born",
            labFormat = labelFormat(suffix = "%")
           )
map12


```

## Percent of Married Families in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of families with married heads of households for each Salt Lake County school district. 

```{r married families, echo=FALSE, message=FALSE}

single <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "B09005", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(single, "acs.colnames")

#convert to data.frame for merging
single_df <- data.frame(paste0(str_pad(single@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(single@geography$schooldistrictunified, 5, "left", pad = "0")),
                         single@estimate[,c("Household Type for Children Under 18 Years in Households (Excluding Householders, Spouses, and Unmarried Partners): Total:", "Household Type for Children Under 18 Years in Households (Excluding Householders, Spouses, and Unmarried Partners): In family households: In married-couple family", "Household Type for Children Under 18 Years in Households (Excluding Householders, Spouses, and Unmarried Partners): In family households: In male householder, no wife present, family", "Household Type for Children Under 18 Years in Households (Excluding Householders, Spouses, and Unmarried Partners): In family households: In female householder, no husband present, family")],
                         stringsAsFactors = FALSE)

single_df <- select(single_df, 1:5)
rownames(single_df)<-1:nrow(single_df)
names(single_df)<-c("GEOID", "totalpop", "married", "singlemale", "singlefemale")
single_df$marriedpercent <- (100* (single_df$married/single_df$totalpop))

##time to merge

single_merged<- geo_join(schooldistrictunified, single_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
single_merged <- single_merged[single_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", single_merged$NAME, "</b><br>", "Percent Family Type Married: ", prettyNum(round(single_merged$marriedpercent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = single_merged$marriedpercent
)

map13<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = single_merged,
              fillColor = ~pal(marriedpercent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = single_merged$marriedpercent,
            position = "bottomright",
            title = "Percent Family Type Married",
            labFormat = labelFormat(suffix = "%")
           )
map13


```

## Percent of Single Families in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of families with single heads of households for each Salt Lake County school district. 

```{r single families, echo=FALSE}
single_df$singlepercent <- (100* ((single_df$singlemale + single_df$singlefemale)/single_df$totalpop))

##time to merge

single_merged<- geo_join(schooldistrictunified, single_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
single_merged <- single_merged[single_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", single_merged$NAME, "</b><br>", "Percent Family Type Single: ", prettyNum(round(single_merged$singlepercent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = single_merged$singlepercent
)

map14<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = single_merged,
              fillColor = ~pal(singlepercent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = single_merged$singlepercent,
            position = "bottomright",
            title = "Percent Family Type Single",
            labFormat = labelFormat(suffix = "%")
           )
map14



```

## Percent Unemployed in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of people unemployed for each Salt Lake County school district. 

```{r unemployment, echo=FALSE}

unemployment <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "B23025", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(unemployment, "acs.colnames")

#convert to data.frame for merging
unemploy_df <- data.frame(paste0(str_pad(unemployment@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(unemployment@geography$schooldistrictunified, 5, "left", pad = "0")),
                         unemployment@estimate[,c("Employment Status for the Population 16 Years and Over: Total:", "Employment Status for the Population 16 Years and Over: In labor force: Civilian labor force: Unemployed")],
                         stringsAsFactors = FALSE)

unemploy_df <- select(unemploy_df, 1:3)
rownames(unemploy_df)<-1:nrow(unemploy_df)
names(unemploy_df)<-c("GEOID", "totalpop", "unemploy")
unemploy_df$percent <- (100* (unemploy_df$unemploy/unemploy_df$totalpop))

##time to merge

unemploy_merged<- geo_join(schooldistrictunified, unemploy_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
unemploy_merged <- unemploy_merged[unemploy_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", unemploy_merged$NAME, "</b><br>", "Percent Unemployed: ", prettyNum(round(unemploy_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = unemploy_merged$percent
)

map16<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = unemploy_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = unemploy_merged$percent,
            position = "bottomright",
            title = "Percent Unemployed",
            labFormat = labelFormat(suffix = "%")
           )
map16


```

## Percent with a Bachelor's Degree and Above in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of people with at least a bachelor's degree for each Salt Lake County school district. 

```{r bachelors degree, echo=FALSE}

bachelor <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "C15003", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(bachelor, "acs.colnames")

#convert to data.frame for merging
bach_df <- data.frame(paste0(str_pad(bachelor@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(bachelor@geography$schooldistrictunified, 5, "left", pad = "0")),
                         bachelor@estimate[,c("Educational Attainment for the Population 25 Years and Over: Total:", "Educational Attainment for the Population 25 Years and Over: Bachelor's degree", "Educational Attainment for the Population 25 Years and Over: Master's degree", "Educational Attainment for the Population 25 Years and Over: Professional school degree", "Educational Attainment for the Population 25 Years and Over: Doctorate degree")],
                         stringsAsFactors = FALSE)

bach_df <- select(bach_df, 1:6)
rownames(bach_df)<-1:nrow(bach_df)
names(bach_df)<-c("GEOID", "totalpop", "bach", "master", "profess", "doc")
bach_df$percent <- (100* ((bach_df$bach + bach_df$master + bach_df$profess + bach_df$doc)/bach_df$totalpop))

##time to merge

bach_merged<- geo_join(schooldistrictunified, bach_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
bach_merged <- bach_merged[bach_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", bach_merged$NAME, "</b><br>", "Percent Bachelor's Degree or Higher: ", prettyNum(round(bach_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = bach_merged$percent
)

map17<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = bach_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = bach_merged$percent,
            position = "bottomright",
            title = "Percent >Bachelor's",
            labFormat = labelFormat(suffix = "%")
           )
map17


```

## Percent of Preschoolers Who Went to Public Preschool in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of preschoolers who went to public preschool for each Salt Lake County school district. 

```{r public school, echo=FALSE}

preschool <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "C14002", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(preschool, "acs.colnames")

#convert to data.frame for merging
preschool_df <- data.frame(paste0(str_pad(preschool@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(preschool@geography$schooldistrictunified, 5, "left", pad = "0")),
                         preschool@estimate[,c("SCHOOL ENROLLMENT BY LEVEL OF SCHOOL BY TYPE OF SCHOOL FOR THE POPULATION 3 YEARS AND OVER: Enrolled in school: Enrolled in nursery school, preschool:", "SCHOOL ENROLLMENT BY LEVEL OF SCHOOL BY TYPE OF SCHOOL FOR THE POPULATION 3 YEARS AND OVER: Enrolled in school: Enrolled in nursery school, preschool: Public school", "SCHOOL ENROLLMENT BY LEVEL OF SCHOOL BY TYPE OF SCHOOL FOR THE POPULATION 3 YEARS AND OVER: Enrolled in school: Enrolled in nursery school, preschool: Private school")],
                         stringsAsFactors = FALSE)

preschool_df <- select(preschool_df, 1:4)
rownames(preschool_df)<-1:nrow(preschool_df)
names(preschool_df)<-c("GEOID", "totalpop", "public", "private")
preschool_df$percent <- (100* ((preschool_df$public)/preschool_df$totalpop))

##time to merge

preschool_merged<- geo_join(schooldistrictunified, preschool_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
preschool_merged <- preschool_merged[preschool_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", preschool_merged$NAME, "</b><br>", "Percent of Kids Who Attended Preschool: Public : ", prettyNum(round(preschool_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = preschool_merged$percent
)

map18<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = preschool_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = preschool_merged$percent,
            position = "bottomright",
            title = "Percent of Kids Who Attended Preschool: Public",
            labFormat = labelFormat(suffix = "%")
           )
map18


```

## Percent of Preschoolers Who Went to Private Preschool in Salt Lake County School Districts 
We used ACS data (from 2015) to map the percent of preschoolers who went to private preschool for each Salt Lake County school district. 

```{r private preschool, echo=FALSE}

preschool_df$percent <- (100* ((preschool_df$private)/preschool_df$totalpop))

##time to merge

preschool_merged<- geo_join(schooldistrictunified, preschool_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
preschool_merged <- preschool_merged[preschool_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", preschool_merged$NAME, "</b><br>", "Percent of Kids Who Attended Preschool: Private : ", prettyNum(round(preschool_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = preschool_merged$percent
)

map19<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = preschool_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = preschool_merged$percent,
            position = "bottomright",
            title = "Percent of Kids Who Attended Preschool: Private",
            labFormat = labelFormat(suffix = "%")
           )
map19


```

## Percent of Families Under the Poverty Line  
We used ACS data (from 2015) to map the percent of under the poverty line for each Salt Lake County school district. 

```{r poverty level, echo=FALSE}

poverty <- acs.fetch(endyear = 2015, span = 1, geography = geo,
                     table.number = "C17026", col.names = "pretty")

#names(attributes(race))
# [1] "endyear"        "span"           "acs.units"      "currency.year"  "modified"       "geography"      "acs.colnames"
# [8] "estimate"       "standard.error" "class"

#attr(poverty, "acs.colnames")

#convert to data.frame for merging
poverty_df <- data.frame(paste0(str_pad(poverty@geography$state, 2, "left", pad="0"),
                                #str_pad(population@geography$county, 3, "left", pad = "0"),
                                str_pad(poverty@geography$schooldistrictunified, 5, "left", pad = "0")),
                         poverty@estimate[,c("Ratio of Income to Poverty Level of Families in the Past 12 Months: Total:", "Ratio of Income to Poverty Level of Families in the Past 12 Months: Under .50", "Ratio of Income to Poverty Level of Families in the Past 12 Months: .50 to .99")],
                         stringsAsFactors = FALSE)

poverty_df <- select(poverty_df, 1:4)
rownames(poverty_df)<-1:nrow(poverty_df)
names(poverty_df)<-c("GEOID", "totalpop", "under50", "under99")
poverty_df$percent <- (100* ((poverty_df$under50 + poverty_df$under99)/poverty_df$totalpop))

##time to merge

poverty_merged<- geo_join(schooldistrictunified, poverty_df, "GEOID", "GEOID", how="inner")
# we need to only include the salt lake county ids
poverty_merged <- poverty_merged[poverty_merged$GEOID %in% c(4900870, 4900420, 4900360, 4900142, 4900600),]
# there are some tracts with no land that we should exclude
# pop_merged <- pop_merged[pop_merged$ALAND>0,]

##pop up makes the box when you click on a specific region, so you can label it with a name or geoid, or something else in your data
popup <- paste0("<b>", poverty_merged$NAME, "</b><br>", "Percent of Families Under the Poverty Level: ", prettyNum(round(poverty_merged$percent,2), big.mark=","))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = poverty_merged$percent
)

map19<-leaflet(qualaddress) %>%
  addMarkers(~long, ~lat, popup = ~as.character(Program.Name), label = ~as.character(Program.Name)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = poverty_merged,
              fillColor = ~pal(percent),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal,
            values = poverty_merged$percent,
            position = "bottomright",
            title = "Percent of Families Under the Poverty Level",
            labFormat = labelFormat(suffix = "%")
           )
map19


```
